Tabular Constraint Learning

### 2016-04-11 Paper/ECAI/tabular.tex
added dependencies into findassignment
updatint the problem statement

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}

@@

\section{Problem Statement}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of Excel constraints and group assignmentsd as follows:

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$\\
    \textbf{Find:}& for each constraint  $c$ in \constraints all the subgroups of $\groups$\\ 
    & satisfying the \CSignature and \CFunction of $c$ \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\section{Approach to Tabular Constraint Learning}
\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Excel constraints, \dependencies -- dependency graph 
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constrainttorder(\constraints,\dependencies)$} 
  \State $v_1, ..., v_n =$ variables of $c$ 
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning \sergey{Samuel, I think \generategroups takes \dependencies as an argument}}
\label{algo:tcl}
\end{algorithm}
In theory each constraint is independent and equally useful. In practice, however, things are different. Assume that for some $X$ and $Y$ it holds that \textit{SUM(X)=Y}, and also we discovered that the vector $Y$ is equal to a vector $Z$. Then, of course, \textit{SUM(X)=Z} holds and it is a valid solution to the sum constraint problem. From a user's perspective the last solution is useless, since he or she already knows that $Y=Z$ and the latter constraint satisfaction follows. To capture this condensed representation \cite{condensed} of solutions, we have introduced a subsumption (generality) order between constraints, which is depicted using \textcolor{red}{dashed} lines in Figure \ref{fig:learning_order}. Besides generality, we know that certain constraints have as a subproblem other constraints, e.g. if $X$ is a \textit{series}, then it is also a \textit{permutation}. If discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all. To capture this solving strategy we introduced a constraint construction order depicted using \textcolor{green}{solid} lines in Figure \ref{fig:learning_order}.

Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.20\textwidth]{figures/constraint_dependency.png}
  \caption{Constraint Learning Order. The \textcolor{green}{solid} arrows indicate a construction order, i.e., learning one constraint is necessary to build the other. The \textcolor{red}{dashed} arrows indicate a subsumption order, the former constraint is used to prune the latter.}
  \label{fig:learning_order}
\end{figure}

@@

\section{Declarative modeling}
\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}

>>>>

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}

@@

A \textit{tabular solution} $S$ to a constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.

\section{Problem Statement}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of Excel constraints and group assignments as follows:

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$\\
    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ \constraints \\ 
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

In theory each constraint is independent and equally useful. In practice, however, things are different. Assume that for some $X$ both \textit{series(X)} and \textit{alldifferent(X)} hold. From a user's perspective the last solution is useless, since he or she already knows that series constraint implies alldifferent constraint. To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}. This matters from implementation perspective as well, if we discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all. A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is a more specific constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.

We call a tabular solution $S$ to a constraint $C$ \textit{the most specific} wrt \dependencies, iff there is no other constraint $C'$ for which $S$ is a solution and $C'$ is more specific in \dependencies.

Let us introduce the version of tabular constraint learning that takes into account the specificity relation between constraints.

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$,\\ 
    & a constraint specificity DAG \dependencies \\
    \textbf{Find:}& all the most specific solution $S$ in \groups\\
    & for each $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

\section{Approach to Tabular Constraint Learning}
\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Excel constraints, \dependencies -- dependency graph 
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constrainttorder(\constraints,\dependencies)$} 
  \State $v_1, ..., v_n =$ variables of $c$ 
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.20\textwidth]{figures/constraint_dependency.png}
  \caption{Constraint Learning Order. The solid arrows indicate a specificity order, i.e., \textit{permutation} $\rightarrow$ \textit{series} means that the series constraint is more specific than permutation. The dashed arrows indicate an inverse construction order, \textit{project} $\dashrightarrow$ \textit{aggregates} means the project constraint is more specific than aggregates. \sergey{Samuel,  can you give an explanation why we keep both arrow and how it is internally motivated?}}
  \label{fig:learning_order}
\end{figure}

@@
